Mapping plant-pollinator interactions across Europe

### 2024-02-12 Manuscript/Ms.qmd
adding idea
2nd paragraph done, now I need something a bit more into detail in the 3rd...

---
title: "Mapping plant-pollinator interactions across Europe"
latex-tinytex: false
format:
  pdf:
    include-in-header: 
      text: |
        \usepackage{lscape}
        \newcommand{\blandscape}{\begin{landscape}}
        \newcommand{\elandscape}{\end{landscape}}
        \usepackage{caption}
        \captionsetup[figure]{labelformat=empty}
bibliography: references.bib
---

@@

1st paragraph

General introduction of how global change impacts plant-pollinator interactions

Maybe expand on some drivers? Climate change, habitat fragmentation, agricultural intensification, urbanization, pollution, pesticides and species' invasions

Highlight the relevance of large scale datasets to understand biological phenomena

2nd paragraph

@@

@@

Introduce research questions

LIST THEM HERE (Main ideas so far)

Questions that we would like to answer:

1\) What are the most common plant a pollinator species? Are those shared across networks? Most common interactions across Europe? Interaction fidelity

2\) Is generalization the rule? Or specialization? How this impacts indirect interactions? Go in the direction of pollinator importance?

@@

@@

>>>>

---
title: "Mapping plant-pollinator interactions across Europe"

format:
  pdf:
    include-in-header: 
      text: |
        \usepackage{lscape}
        \newcommand{\blandscape}{\begin{landscape}}
        \newcommand{\elandscape}{\end{landscape}}
        \usepackage{caption}
        \captionsetup[figure]{labelformat=empty}
bibliography: references.bib
latex-tinytex: false
link-citations: TRUE
linkcolor: RoyalBlue
urlcolor: RoyalBlue
---

@@

The interaction between plants and pollinators is one of the most well-documented mutualisms on Earth. Plant-pollinator interactions involve a great diversity of species, largely attributed to their coevolutive history [@ollerton2017], and are critically important for terrestrial biodiversity and the human economy. The synergistic effects of climate change with other global change pressures are threatening worldwide biodiversity [@sala2000; @bellard2014], including plant and pollinator species as well as their interactions [@gonzalez2013; @settele2016; @eichenberg2021]. Under this scenario, the increasing availability of biodiversity data plays a major role in our ecological understanding of species and guiding conservation planning [@heberling2021]. However, our knowledge of plant and pollinator species and their network of interactions still exhibits major temporal, spatial and taxonomic biases [@archer2014; @troia2016; @poisot2021; @marshall2024] that limit our ability to effectively protect their biodiversity.

Europe is one of the continents with a larger amount of available biodiversity data [@proencca2017], yet still exhibits major gaps [@wetzel2018; @bennett2018]. While species checklists need to be treated carefully, especially at a macro-ecological scale [@grenie2023], the growing number of European plant and pollinator checklists along with occurrence data, is setting a foundation for the conservation of its flora and their pollinators. However, species richness is just one component of biodiversity and documenting the interaction between plants and pollinators is essential for understanding the fate of the species within an ecosystem [@jordano2016]. Numerous works have studied plant-pollinator interactions in the last decades, originating thousands of plant-pollinator interaction networks worldwide. Several initiatives have tried to integrate plant-pollinator interaction data into databases such as *Mangal* [@poisot2016] or *GloBI* [@poelen2014], resulting in numerous large scale comparative analyses that have enhanced our understanding of the ecology of plants and pollinators [e.g., European wild bee data trends\; @marshall2024]. A major drawback of these global databases is the limited number of studies at regional scales and the lack of additional information (e.g., sampling effort or methodology), which hampers our understanding of plant-pollinator interactions. Thus, in a scenario of rapid environmental change, integrating robust informative plant-pollinator interaction data is essential for evaluating our current knowledge at the European level, guiding conservation planning, and establishing a solid foundation for global change research.

@@

Low connectance?

Species occurrences

Threatened species? Red lists?

Discuss actual TRENDS!

@@

Here, we present the SafeNet database, an acronym derived from the European project Safeguard and plant-pollinator networks, which contains harmonize interaction data of plants and pollinators at European level.

@@

Define study area!

@@

Contextualise European diversity of plant and pollinator species with global biodiveristy
